---
title: "ETC5513 Project"
author:
- familyname: XXX
  othernames: XXX
  address: Monash University
  email: XXX
  correspondingauthor: true
  qualifications:  XXX
- familyname: Nguyen
  othernames: James Louis
  address: Monash University
  email: jlngu6@student.monash.edu
  correspondingauthor: true
  qualifications:  XXX
- familyname: Effendy
  othernames: Jimmy
  address: Monash University
  email: jeff0001@student.monash.edu
  correspondingauthor: true
  qualifications:  B.Com
- familyname: XXX
  othernames: XXX
  address: Monash University
  email: XXX
  correspondingauthor: true
  qualifications:  XXX
department: Econometrics and Business Statistics \newline add names &\newline add names
organization: Monash University
bibliography: references.bib
biblio-style: authoryear-comp
linestretch: 1.5
output:
  bookdown::pdf_document2:
    template: monashreport2.tex
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
    toc: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r libraries, include=FALSE}
library(tidyverse)
library(kableExtra)
library(bookdown)
library(lubridate)
```


```{r data-read, include=FALSE}
data = read.csv("data/wdi_data.csv")
data = data[1:35,]

names(data)[5:54] = c(1970:2019)
names(data)[3] = c("Series")

data_value = data %>%
  select('1970':'2019')

data_value = data.frame(lapply(data_value, as.character), stringsAsFactors =FALSE)
data_value = data.frame(lapply(data_value, as.numeric))

data = data.frame(Series = data$Series, data_value)
names(data)[2:51] = c(1970:2019)

data_long = data %>%
  pivot_longer(`1970`:`2019`, names_to = "Year") %>%
  mutate(Year = as.factor(Year))

#use data_long for analysis.
```


\section*{Introduction}

\section*{Motivation}

\section*{}

\section*{}

\section*{Education}

```{r education-expenditure}
data_education <- data_long %>% 
  filter(Series %in% (c("Ratio of female to male labor force participation rate (%) (national estimate)",
                        "Adjusted savings: education expenditure (current US$)",
                        "Government expenditure on education, total (% of GDP)",
                        "Adjusted savings: education expenditure (% of GNI)")))

#adding decades information to the dataframe
data_education$Year <- as.character(data_education$Year)
data_education$Year <- as.numeric(data_education$Year)

data_education <- data_education %>% mutate(decade = floor(Year/10)*10)

```

```{r ratio-plot-decade, fig.cap="Ratio of female to male labor force participation rate (%) (national estimate) by decades"}

data_education %>% filter(Series == "Ratio of female to male labor force participation rate (%) (national estimate)") %>%
  group_by(decade) %>% 
  summarise(average = mean(value, na.rm = TRUE)) %>% 
  ggplot(aes(x = decade,
             y = average)) +
  geom_col()
```

In the past 50 years, women's participation in the work force has increased steadily [@stier2000women].

It is believed that gender inequality in education is a major contributing factor in gender inequalities in adulthood, especially in the labour market [@marks2008accounting].

In this section, the report is going to examine the state of education in Israel and whether there is an association between education and the increase of women's employment in Israel.

\subsection*{Education Spending in Israel}

```{r read-edu-pop}
library(simputation)

read_edu_pop <- read.csv("data/adult_edu_pop.csv") %>% 
  rename(location = `ï..LOCATION`)
read_edu_spending <- read.csv("data/edu_spending.csv") %>%
  rename(location = `ï..LOCATION`)
read_pop_tert <- read.csv("data/pop_tert_edu.csv") %>% 
  rename(location = `ï..LOCATION`)

#wrangling read_edu_pop dataset
data_edu_pop <- read_edu_pop %>% 
  filter(location == "ISR") %>% 
  mutate(SUBJECT = case_when(
    SUBJECT == "BUPPSRY" ~ "below_upper_secondary",
    SUBJECT == "TRY" ~ "tertiary",
    SUBJECT == "UPPSRY" ~ "upper_secondary",
    SUBJECT == "TRY_MEN" ~ "tertiary_men",
    SUBJECT == "TRY_WOMEN" ~ "tertiary_women",
    SUBJECT == "UPPSRY_MEN" ~ "upper_secondary_men",
    SUBJECT == "UPPSRY_WOMEN" ~ "upper_secondary_women",
    TRUE ~ as.character(SUBJECT)
  ))

#wrangling read_edu_spending dataset
data_edu_spending <- read_edu_spending %>%
  filter(location == "ISR",
         MEASURE == "USD_STUDENT") %>%
  select(TIME, SUBJECT, Value) %>% 
  mutate(SUBJECT = case_when(
    SUBJECT == "SRY" ~ "secondary",
    SUBJECT == "TRY" ~ "tertiary",
    SUBJECT == "PRY_NTRY" ~ "primary_to_post_secondary_non_tertiary",
    SUBJECT == "EARLYCHILDEDU" ~ "early_childhood_education",
    SUBJECT == "PRY" ~ "primary",
    TRUE ~ as.character(SUBJECT)
  )) %>% 
  filter(TIME >= 2008)

#impute missing data in data_edu_spending
data_edu_spending_wide <- data_edu_spending %>% 
  pivot_wider(names_from = SUBJECT,
              values_from = Value) %>% 
  select(-secondary) %>% 
  filter(TIME != 2016) %>% 
  rename(time = TIME)

data_edu_spending_wide_imp <- data_edu_spending_wide %>% 
  as.data.frame() %>% 
  impute_lm(formula = primary_to_post_secondary_non_tertiary ~ time) %>% 
  impute_lm(formula = early_childhood_education ~ time)

data_edu_spending_long_imp <- data_edu_spending_wide_imp %>% 
  pivot_longer(cols = -time,
               names_to = "subject",
               values_to = "value")
```

```{r total-edu-spending, fig.cap="Yearly Average Education Expenditure (US$) in Israel by Decades"}
data_education %>% filter(Series == "Adjusted savings: education expenditure (current US$)") %>%
  #summarising education expenditure by decades
  group_by(decade) %>%
  summarise(avg_spending = mean(value, na.rm = TRUE)) %>% 
  ggplot(aes(x = decade,
             y = avg_spending)) +
  geom_col()
```
The education spending spent by the Israeli government is shown in figure \@ref(fig:total-edu-spending).

The average of funds invested by the government doubled the size almost every decade.

It started with yearly average of 834 million US dollars in 1970s, and it reached 15 billion US dollars in 2010s.

```{r data-edu-spend, fig.cap="Education Spending in Israel"}
data_edu_spending_long_imp %>%
  filter(subject != "primary") %>% 
  ggplot(aes(x = time,
             y = value,
             color = subject)) +
  geom_point() +
  geom_line() +
  theme(legend.position = "bottom") +
  labs(x = "Year", y = "Spending (US$)")
```

Figure \@ref(fig:data-edu-spend) reflects how the Israeli government invested the education spending from 2008 to 2015.

The figure shows that the government spent the majority of the funds on tertiary educations.

It is then followed by primary to post secondary (non-teritary) and early childhood education respectively.

```{r edu-pop-female, fig.cap= "Female Adult Education Level of 25-64 Year-olds in Israel"}
data_edu_pop %>% filter(SUBJECT %in% c("tertiary_women",
                                       "upper_secondary_women")) %>% 
  ggplot(aes(x = TIME,
             y = Value,
             color = SUBJECT)) +
  geom_point() +
  geom_line() +
  theme(legend.position = "bottom") +
  labs(x = "Year", y = "% of 25-64 year-olds")
```
\subsection*{Adult Education Level in Israel}



\subsection*{Education and Female Participation in Labour Force}

OECD 2016:
* Tertiary education has a strong impact in the labour market

* Tertiary-educated adults in Israel have a relatively high employment rate (86%), which reaches its peak of 92%
among those who attained a doctoral or equivalent degree.

* Most of those enrolled in tertiary programmes in the academic year 2013/14 attend private institutions (83%),
the third highest private enrolment rate of all OECD countries. Most are enrolled in government-dependent
private institutions (66%); 17% are enrolled in independent private institutions. 



We also note the synergetic interaction between higher education and increased labor force participation: A major reason for this rise is that employment opportunities increase with higher education and that these offer potentially higher financial remuneration, increasing the incentives to enter the labor market. Furthermore, the higher wages that accompany higher education levels make it increasingly possible to afford the costs of domestic help and childcare, thus making entry into the work force more economically feasible and attractive.

```{r labor-edu-scatterplot, fig.cap = "Scatter Plot between Ratio of Female to Male Labour Force Participation Rate and Education Expenditure"}

data_education %>% filter(Series %in% (c("Adjusted savings: education expenditure (current US$)",
                                        "Ratio of female to male labor force participation rate (%) (national estimate)"))) %>%
  pivot_wider(names_from = Series,
              values_from = value) %>% 
  ggplot(aes(x = `Adjusted savings: education expenditure (current US$)`,
             y = `Ratio of female to male labor force participation rate (%) (national estimate)`,
             label = Year)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  theme(aspect.ratio = 1)

```

\section*{}

\section*{Model}

\section*{Conclusion}