---
title: "ETC5513 Project"
author:
- familyname: XXX
  othernames: XXX
  address: Monash University
  email: XXX
  correspondingauthor: true
  qualifications:  XXX
- familyname: Nguyen
  othernames: James Louis
  address: Monash University
  email: jlngu6@student.monash.edu
  correspondingauthor: true
  qualifications:  XXX
- familyname: Lyubic
  othernames: Samuel
  address: Monash University
  email: slyu0007@student.monash.edu
  correspondingauthor: true
  qualifications:  XXX
- familyname: Tirodkar
  othernames: Siddhant
  address: Monash University
  email: stir0003@student.monash.edu
  correspondingauthor: true
  qualifications:  XXX
department: Econometrics and Business Statistics \newline add names &\newline add names
organization: Monash University
bibliography: references.bib
biblio-style: authoryear-comp
linestretch: 1.5
output:
  bookdown::pdf_document2:
    template: monashreport2.tex
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
    toc: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, include=FALSE}
library(tidyverse)
library(kableExtra)
library(bookdown)
library(lubridate)
library(knitr)
library(gridExtra)
library(imputeTS)
library(gridExtra)
library(grid)
```


```{r data-read, include=FALSE}
data = read.csv("data/wdi_data.csv")
data = data[1:36,]

names(data)[5:54] = c(1970:2019)
names(data)[3] = c("Series")

data_value = data %>%
  select('1970':'2019')

data_value = data.frame(lapply(data_value, as.character), stringsAsFactors =FALSE)
data_value = data.frame(lapply(data_value, as.numeric))

data = data.frame(Series = data$Series, data_value)
names(data)[2:51] = c(1970:2019)

data_long = data %>%
  pivot_longer(`1970`:`2019`, names_to = "Year") %>%
  mutate(Year = as.factor(Year))

#use data_long for analysis.
```


\section*{Introduction}

\section*{Motivation}

\section*{}

\section*{Analysing energy consumption and renewable sources}

Israel has seen significant changes in the distribution of genders in the labour force. Though many direct effects can be explored, this section looks at changes in attitude by analysing Israel's move to more renewable sources of energy. There may be a link between the a transition to change in political thinking in priority. This may also be attributed to a general change in attitude, and could provide a reasonable proxy for our model.

For this section of the report, we will be looking at World Bank Sustainable for All data (@TheWorldBank2018). This data set comprises of electricity data between 1990 and 2015. 
```{r include=FALSE}
data_sea = read.csv("data/SEA_data.csv")
names(data_sea)=c("Country Name", "Country Code", "Year", "Year Code",
                  "Renewable electricity share of total electricity output (%)",
                  "Renewable energy consumption (TJ)",
                  "Total final energy consumption (TFEC) (TJ)",
                  "Renewable electricity output (GWh)",
                  "Renewable energy share of TFEC (%)")
data_sea = data_sea[1:26,]

data_sea_long = data_sea %>%
  pivot_longer(`Renewable electricity share of total electricity output (%)`:`Renewable energy share of TFEC (%)`,names_to = "Series") %>%
  mutate(value = as.character(value)) %>%
  mutate(value = as.numeric(value))
```


```{r compare-renewable-ftm, fig.cap="Comparing female to male ratio to renewable electricity output share", echo=FALSE}
p1 = data_sea_long %>% 
  filter(Series == "Renewable electricity share of total electricity output (%)") %>%
  ggplot(aes(x = Year, y = value)) + 
  geom_line(aes(group = 1)) + 
    theme(axis.text.x = element_text(angle = 90),
        axis.title.x = element_blank()) + 
  ylab("Renewable electricity ratio (%)") 

p2 = data_long %>%
  filter(Series == "Ratio of female to male labor force participation rate (%) (national estimate)",
         Year %in% c(1990:2015)) %>%
  mutate(Year = as.character(Year)) %>%
  mutate(Year = as.numeric(Year)) %>%
  ggplot(aes(x = Year, y=value)) + 
  geom_line(aes(group=1)) + 
  theme(axis.text.x = element_text(angle = 90),
        axis.title.x = element_blank()) + 
  ylab("Female to male labor force (%)")

grid.arrange(p1,p2, ncol=2,
             top=textGrob("Comparing female to male (%) and Renewable electricity share output (%)",gp=gpar(fontsize=10,font=1)),
             bottom=textGrob("Year",gp=gpar(fontsize=10,font=1)))
```

Comparing the ratio of female to male labor force participation rates to renewable electricity share in Figure \@ref(fig:compare-renewable-ftm), we note a more consistent growth of the ratio over time, whereas the renewable electricity ratio stays consistent until 2007, after which we observe a sharp increase. To understand this increase, we must delve more into the renewables future plan for Israel. Established in 2016 were steps to reach a target of 17% Renewable Energy production by 2030, with interim targets of 10% and 13% by 2020/2023 respectively (@IEA2019). Unfortunately, the data set used for this analysis does not have data past 2015, and as a result, we cannot identify the effect which the announcement could have made.

Looking at the change in electricity share in Figure \@ref(fig:renewable-compare):

```{r renewable-compare, fig.cap="Looking at Renewable electricity share, and difference change", echo=FALSE}
data_sea_long_1 = data_sea_long %>%
  filter(Series == "Renewable electricity share of total electricity output (%)")

data_sea_long_1 = cbind(data_sea_long_1, difference = c(NA, diff(data_sea_long_1$value))) %>%
  drop_na()


p3 = data_sea_long %>% 
  filter(Series == "Renewable electricity share of total electricity output (%)") %>%
  ggplot(aes(x = Year, y = value)) + 
  geom_line(aes(group = 1)) + 
    theme(axis.text.x = element_text(angle = 90),
        axis.title.y = element_blank(),
        axis.title.x = element_blank()) + 
  ylab("Renewable electricity share of total electricity output (%)")


p4 = data_sea_long_1 %>% 
  drop_na() %>%
  ggplot(aes(x=Year, y=difference)) +
  geom_line(aes(group=1)) + ylab("Difference change (%)")

grid.arrange(p3,p4, top=textGrob("Renewable electricity share of total electricity output (%)",gp=gpar(fontsize=15,font=1)))
```

We can see that the difference seems to centre around zero prior to 2007, as was noted previously, however no clear perfect correlation can be seen when comparing the differences in Figure \@ref(fig:renewable-compare) and Figure \@ref(fig:compare-renewable-ftm). 



```{r table-setup, include=FALSE}
mean_difference = mean(data_sea_long_1[["difference"]])

table_forecast = data.frame(Year = data_sea_long_1$Year, 
                            Value = data_sea_long_1$value)

val = c(table_forecast$Value[25] + seq(1:5)*mean_difference)

fc = data.frame(Year = c(2016:2020), 
                Value = val)

table_forecast = rbind(table_forecast, fc)

names(table_forecast)[2] = c("Renewable electricity share of total electricity output (%)")
```

\pagebreak

```{r fcrenew, echo=FALSE}
kable(tail(table_forecast,5), booktabs = TRUE,
      caption = "Forecast of renewable electricity share") %>%
  kable_styling(full_width = FALSE, "striped", position = "center", font_size = 10)
```

Table \@ref(tab:fcrenew) shows a forecast of the renewable electricity share of total electricity output in the next five years. This was conducted using a mean forecasting method. If current policies were followed, the target set in 2016 may not be met.

Perhaps, renewable energy is not exactly suitable in representing some change in attitude, Burke (@Burke2018) and Sung (@Sung2018)  finds that renewable transition for the energy industry has been promoted by governments and markets, whereas traditional energy companies tend to hamper this progress. they also find that the general public do not have directly influence the transition, but indirectly affects it through the markets and governments. It may be suitable for the ratio of renewable energy output to proxy the strength of effect that the government/market has, and by extension, how the public affects decision making in the government.

Let's look at total energy consumed in relation to the female to male labour force ratio: 

```{r energy-ftm, fig.cap = "Comparing total energy consumption and female to male ratio.", echo=FALSE}
p5 = data_sea_long %>% 
  filter(Series %in% c("Total final energy consumption (TFEC) (TJ)")) %>%
  ggplot(aes(x = Year, y = value)) + 
  geom_line() + 
    theme(axis.text.x = element_text(angle = 90),
          axis.title.x = element_blank()) +
  ylab("Total final energy consumption (TFEC) (TJ)")

grid.arrange(p5,p2, ncol=2,
             top=textGrob("Comparing total energy consumed (TJ) and female to male (%)",gp=gpar(fontsize=10,font=1)),
             bottom=textGrob("Year",gp=gpar(fontsize=10,font=1)))
```

We can see a similar trend between series in Figure \@ref(fig:energy-ftm). We could use total energy consumption as a way to proxy general changes in attitudes. Andrae (@Andrae2015) discusses the general increase in electricity usage due to evolving communication technologies, including consumer devices. Here we note a significant increase in energy consumption over time, most likely attributable to an increased reliance on technology, including information technology. With a access to wider information, it is reasonable to expect some sort of changing attiude. 

To conclude this section, we can use renewable electricity output ratio to directly model some government/market influence, and indirectly people. It may also be more reasonable to use total energy consumption to model changing attitudes.

\section*{Male and Female Mortality Rates and Birth Rate}

With Israel as our chosen country, this section will be analyzing the relationship between our repose variable, _Ratio of female to male labor force participation rate (%) (national estimate)_, and the following economic indicators: _Birth rate, crude (per 1,000 people)_, _Mortality rate, adult, female (per 1,000 female adults)_ and _Mortality rate, adult, male (per 1,000 male adults)_ to assess their impact on the change of the female to male labor force ratio to ultimately address the change in how the gender stereotype is viewed in Israel. 
Across history, the traditional stereotype on gender roles has been one that depicts males roles as the provider and women as the caretaker and child bearer resulting in the notion that a woman's role is at home which has led to less occupational opportunities and aspirations @GenderRoles. These stereotypes are further perpetuated through religion, as seen by the two most present religion in Israel - Judaism and Islam where the traditional identity for involves leadership for men and domesticity for females @Woodhead in all facets of life from work, private, public and in the army @ServiceIDF. However, it does seem as though this traditional view is being confronted and dismantled over time with more religious and non religious Israeli woman taking up education and being accepted in a range of occupational opportunities and army duties @Levy2006, as such birth rate and male and female mortality rates are prime variables to potentially explain the change in the labor force ratio thus indicating the potential change in how gender stereotypes are viewed. 

The variables for this section are isolated into their own data set in both long and wide form.
```{r create-long-wide-objects-with-required-variables}
mortality_birth_rate_long <- data_long %>%
  filter(Series %in% c("Ratio of female to male labor force participation rate (%) (national estimate)", "Mortality rate, adult, female (per 1,000 female adults)", "Mortality rate, adult, male (per 1,000 male adults)", "Birth rate, crude (per 1,000 people)")) %>%
  arrange(Year) %>%
  slice(65:188)
```

### The variable change overtime

```{r figure-female-and-male-mortality-and-birth-rate-and-labor-force-ratio, include = FALSE}
figure_female_male_mortality <- mortality_birth_rate_long %>%
  filter(Series %in% c("Mortality rate, adult, female (per 1,000 female adults)", 
                       "Mortality rate, adult, male (per 1,000 male adults)")) %>%
  ggplot(aes(x = Year,
             y = value,
             group = Series,
             color = Series)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "right") +
  theme_bw() +
  labs(title = "Male and female mortality rate, per 1000 adults",
       y = "Mortality rate per 1000 people",
       x = "Year") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) 

figure_birth_rate <- mortality_birth_rate_long %>%
  filter(Series == "Birth rate, crude (per 1,000 people)") %>%
  ggplot(aes(x = Year,
             y = value,
             group = Series,
             color = Series)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "right") +
  theme_bw() +
  labs(title = "Birth rate per 1000 people",
       y = "Birth rate, crude (per 1,000 people)",
       x = "Year") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

figure_labor_force_ratio <- mortality_birth_rate_long %>%
  filter(Series == "Ratio of female to male labor force participation rate (%) (national estimate)") %>%
  ggplot(aes(x = Year,
             y = value,
             group = Series,
             color = Series)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "right") +
  theme_bw() +
  labs(title = "Ratio female to male labor force participation",
       y = "Female to male ratio %",
       x = "Year") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) 
  
```

```{r combining-the-variable-figures, fig.cap = "Each figure visualises the chosen variables", fig.width = 10, fig.height = 8}
figure_combined_birth_ratio_mortality <- grid.arrange(figure_labor_force_ratio, figure_birth_rate, figure_female_male_mortality)
```

```{r percent-change-over-the-observed-time-period, fig.cap = "Figure showing the percentage chance in each of the chosen variables"}
pct_change_mortality_variables <- mortality_birth_rate_long %>%
  filter(Year %in% c("1986", "2016")) %>% 
  pivot_wider(names_from = Year,
              values_from = value) %>%
  mutate(percent_change_1986_to_2016 = ((`2016`-`1986`)/`1986`)*100) %>%
  kable(caption = "The total percentage change from 1986 to 2016, per each variable") %>%
  kable_styling(bootstrap_options = c("hover", "striped"),
                latex_options = "hold_position") 
pct_change_mortality_variables    
```

Figure \@ref(fig:combining-the-variable-figures) visualizes the change over time for each of the chosen variables in this section, with the selected variable on the y axis and year along the x axis. 

Key takeaways:
  - The labor force participation figure shows that the ratio of female to male labor force participation rate has increased from 1986 to 2016, by 41.01% as shown by Table \@ref(tab:pct_change_mortality_variables).
  - The birth rate figure shows the birth rate per 1000 people overall shows a reduction in the birth rate has increased from 1986 to 2016, by -8.23% as shown by Table \@ref(tab:pct_change_mortality_variables).
  - The mortality rate table shows the mortality rate for females per 1000 female adults in orange and the mortality rate for males per 1000 males in blue. Both indicators have reduced from 1986 to 2016 at a relatively even rate, with female mortality reducing by -48.94% and male mortality reducing by -45.80%, as shown by Table \@ref(tab:pct_change_mortality_variables).

### Correlation comparison

The explanatory variables are presented as ratios and in order for their results to be relative to the explanatory variable and to allow for analysis to take place, they are turned into percentages. The new column percent_value shows the percentage of birth rate and female and male mortality rates for every 1000 people, from which a new data set will be created showing the percentage results for all the variables in this section.
```{r ratios-turned-into-percentages, include = FALSE}
variables_turned_into_percentage <- mortality_birth_rate_long %>%
  filter(!Series == "Ratio of female to male labor force participation rate (%) (national estimate)") %>%
  mutate(percent_value = ((value/1000)*100))
```

```{r combine-percent-results-response-variable-and-explanatory, include = FALSE}
labor_force_ratio <- mortality_birth_rate_long %>%
  filter(Series == "Ratio of female to male labor force participation rate (%) (national estimate)") %>%
  mutate(percent_value = value)
percentage_results_all_variables <- rbind(variables_turned_into_percentage, labor_force_ratio)
```

```{r pivot-wider-percentage-results, include = FALSE}
percentage_results_wider <- percentage_results_all_variables %>% 
  select(Series, Year, percent_value) %>%
  pivot_wider(names_from = Series,
              values_from = percent_value) %>%  
  rename(Pct_male_mortality_rate_per_1000_male_adults = "Mortality rate, adult, male (per 1,000 male adults)",
         Pct_Female_mortality_rate_1000_female_adults = "Mortality rate, adult, female (per 1,000 female adults)",
         Pct_Birth_rate_per_1000_people = "Birth rate, crude (per 1,000 people)",
         Labor_force_rate = "Ratio of female to male labor force participation rate (%) (national estimate)")
```

```{r correlation-between-response-explanatory, fig.cap = "The table shows the correlation coefficient between each explanatory variable and the response variable"}
correlation_results_male_mort <- cor(percentage_results_wider$Pct_male_mortality_rate_per_1000_male_adults, percentage_results_wider$Labor_force_rate)
correlation_results_female_mort <- cor(percentage_results_wider$Pct_Female_mortality_rate_1000_female_adults, percentage_results_wider$Labor_force_rate) 
correlation_results_birth_rate <- cor(percentage_results_wider$Pct_Birth_rate_per_1000_people, percentage_results_wider$Labor_force_rate)

table_correlation <- tibble(correlation_results_male_mort, correlation_results_female_mort, correlation_results_birth_rate) %>%
  kable(caption = "The correlation between the percentage results, per 1000 people, of the response variables and the ratio of female to male labor force participation rate (%)") %>%
  kable_styling(bootstrap_options = c("hover", "striped"),
                latex_options = "hold_position") 
table_correlation
```

```{r figure-correlation-female-mort-labor-force, fig.cap = "The figure displays the relationship between the percentage results of female mortality rates per 1000 adult females and the Ratio of female to male labor force participation rate (%) (national estimate)"}
correlation_pct_results_female_mort <- percentage_results_wider %>%
  ggplot(aes(x = Labor_force_rate, 
             y = Pct_Female_mortality_rate_1000_female_adults)) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE) +
  labs(title = "Correlation between the female mortality 
                rate and labor force ratio",
       y = "Female mortality per 1000 adult females, as a percentage",
       x = "Ratio female to male labor force participation")
correlation_pct_results_female_mort
```

```{r figure-correlation-male-mort-labor-force, fig.cap = "The figure displays the relationship between the percentage results of male mortality rates per 1000 adult females and the Ratio of female to male labor force participation rate (%) (national estimate)"}
correlation_pct_results_male_mort <- percentage_results_wider %>%
  ggplot(aes(x = Labor_force_rate, 
             y = Pct_male_mortality_rate_per_1000_male_adults)) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE) +
  labs(title = "Correlation between the male mortality 
                rate and labor force ratio",
       y = "Male mortality per 1000 adult males, as a percentage",
       x = "Ratio female to male labor force participation")
correlation_pct_results_male_mort
```

```{r figure-correlation-birth-rate-labor-force, fig.cap = "The figure displays the relationship between the percentage results of birth rates per 1000 adult females and the Ratio of female to male labor force participation rate (%) (national estimate)", echo = FALSE}
correlation_pct_results_birth_rate <- percentage_results_wider %>%
  ggplot(aes(x = Labor_force_rate, 
             y = Pct_Birth_rate_per_1000_people)) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE) +
  labs(title = "Correlation between birth rate and labor force ratio",
       y = "Birth rate per 1000 people, as a percentage",
       x = "Ratio female to male labor force participation")
correlation_pct_results_birth_rate
```

Figure \@ref(fig:figure-correlation-birth-rate-labor-force) visualizes the correlation between birth rate and the female to male labor force ratio, with birth rate, crude (per 1,000 people) on the y axis and the ratio of female to male labor force participation rate on the x axis. The figure shows a negative correlation between the two variables however, there are some scattered results where both birth rate and female to male labor force ratio are both relatively low which is depicted by the correlation coefficient of -0.60 as shown by Table \@ref(tab:correlation-between-response-explanatory).


Figure \@ref(fig:figure-correlation-male-mort-labor-force) visualizes the correlation between male mortality rate and the female to male labor force ratio, with male mortality rate per 1,000 male adults on the y axis and the ratio of female to male labor force participation rate on the x axis. The figure displays a strong negative correlation between the two variables which is backed up by the correlation coefficient of -.96 in Table \@ref(tab:correlation-between-response-explanatory). 

Figure \@ref(fig:figure-correlation-female-mort-labor-force) visualizes the correlation between female mortality rate and the female to male labor force ratio, with female mortality rate per 1,000 female adults on the y axis and the ratio of female to male labor force participation rate on the x axis. The figure displays a strong negative correlation between the two variables which is backed up by the correlation coefficient of -.99 in Table \@ref(tab:correlation-between-response-explanatory).

As evident by the analysis conducted in this section, there is a clear positive correlation between the three explanatory variables and the the ratio of female to male labor force participation rate and as such can work as indicator for the change in view of gender stereotypes in Israel.  



\section*{Labour Force Analysis}

# Overview

This section of the report will focus on analysing various aspects of the Israeli labour force with respect to the research question that we have chosen which is "How have the views towards gender stereotypes changed over time in Israel?". The Israeli labour force has come a long way from being male dominated to being almost equally divided between men and women. The transition of the Israeli labour force shall be explored through the years 1991-2019 in this section of the report.
```{r labour-data}
israel_labour <- data_long %>% dplyr::filter(Series %in% c("Labor force, female (% of total labor force)",
                                                      "Ratio of female to male labor force participation rate (%) (national estimate)", "Self-employed, female (% of female employment) (modeled ILO estimate)" , "Wage and salaried workers, female (% of female employment) (modeled ILO estimate)",
                                                      "Labor force participation rate, female (% of female population ages 15-64) (modeled ILO estimate)","Labor force participation rate, male (% of male population ages 15-64) (modeled ILO estimate)"),
                                        Year %in% c(1991:2019))

```

```{r showcase}
israel_display <- israel_labour %>% pivot_wider(names_from = Series, values_from = value)

  kable(head(israel_display), caption = "Variables Summary") %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                latex_options = "hold_position") %>%
    column_spec(1, width = "2em") %>%
    column_spec(2, width = "5em") %>%
    column_spec(3, width = "5em") %>%
    column_spec(4, width = "5em") %>%
    column_spec(5, width = "5em") %>%
    column_spec(6, width = "5em") %>%
    column_spec(7, width = "5em")
    
  

```

Table \@ref(tab:showcase) exhibits the variables that have been used in this section of the report and displays the first 5 rows. The ratio of self-employed females and female labour force participation are the explanatory variable and the ratio of female to male participation rate is the response variable.The table gives a brief overview of how the variables change from the period 1991-1996.

```{r change}
israel_labour$Year <-  as.numeric(as.character(israel_labour$Year))

```

```{r graph1,fig.cap="% Female labour force out of total labour force"}

plot_1 <- israel_labour %>% dplyr::filter(Series == "Labor force, female (% of total labor force)") %>%
  ggplot(aes(x = Year, y = value, color = Series))+
  geom_line()+
  xlab("Year")+
  ylab("Total Labour force(%)")+
  theme(legend.position="bottom")
  

plot_1
```

Figure \@ref(fig:graph1) depicts the gentle rise in the female labour force which was around 40% of the total labour force in 1990 to around 47% around 2020. One of the main reasons behind the surge in female participation is due to the educational attainment of women in Israel. Also, in 2006 and 2007 numerous amendments were made to women employment law to protect the rights of women at workplace and provide equal employment opportunities which has proved to be pivotal in increasing female participation in the labour force. [Source](https://mfa.gov.il/MFA/AboutIsrael/State/Law/Pages/Ensuring-equal-rights-for-women-in-Israel.aspx)

# Self-employed v/s wage & salaried women

```{r graph2,fig.cap="Self-employed v/s Wage & salaried female workers"}

plot_2 <- israel_labour %>% dplyr::filter(Series %in% c("Self-employed, female (% of female employment) (modeled ILO estimate)" , "Wage and salaried workers, female (% of female employment) (modeled ILO estimate)")) %>% 
  ggplot(aes(x = Year, y = value, fill = Series))+
  geom_col()+
  xlab("Year")+
  ylab("Total female employment (%)")+
  theme(legend.position="bottom")
  

plot_2
```

Figure \@ref(fig:graph2) shows the slow but gradual increase in the percentage of self-employed females in the female labour workforce. The female labour workforce comprises of self-employed women and wage and salaried women. The percentage of self-employed women in Israel has increased from 5.9 to 9.2 in 3 decades. The gradual change can be attributed to strong family orientation present in the Israeli culture and the fact that a working mother has to always give priorities to family responsibilities. But the main contributing factor to the increase was achievement motivations and economic necessities more than educational attainment or previous entrepreneurial experience. @lerner1997israeli

# Female v/s Male labour force participation

```{r graph3,fig.cap="Female v/s Male labour force participation"}

plot_3 <- israel_labour %>% dplyr::filter(Series %in% c("Labor force participation rate, female (% of female population ages 15-64) (modeled ILO estimate)","Labor force participation rate, male (% of male population ages 15-64) (modeled ILO estimate)")) %>%
  ggplot(aes(x=Year,y=value, color = Series))+
  geom_line()+
  xlab("Year")+
  ylab("Labour force participation rate")+
  theme(legend.position="bottom")

plot_3
```

Figure \@ref(fig:graph3) compares the female v/s male labour force participation amongst the total population in the age group of 15-64 females and males respectively. It depicts the percentage of the female population that is engaged in the labour force in the working age population. This is a very good representation of the increased number of participation of women in the labour force which has increased almost 20% in the past 30 years which authenticates that views towards gender stereotypes are changing over time in Israel. This shows that Israeli women are more willing and motivated to be involved in the labour force than they ever were. 

# Female labour force v/s female-male participation rate
```{r graph4}
plot_4 <- israel_labour %>% dplyr::filter(Series %in% c("Ratio of female to male labor force participation rate (%) (national estimate)")) %>%
  ggplot(aes(x=Year,y=value, color = Series))+
  geom_line()+
  xlab("Year")+
  ylab("Female to male Partcipation rate(% of total)")+
  theme(legend.position="bottom")
                                         
```

```{r graph5,fig.cap="Comparing female labour force with female-male participation rate"}

grid.arrange(plot_1,plot_4, ncol=2)

```

Figure \@ref(fig:graph5) verifies that the percentage of females in the total labour force is positively correlated with the female to male labour force participation rate as both show a positive trend over time and have increased significantly since 1991 to 2019.

\section*{Model}

\section*{Conclusion}

references