---
title: "ETC5513 Project"
author:
- familyname: XXX
  othernames: XXX
  address: Monash University
  email: XXX
  correspondingauthor: true
  qualifications:  XXX
- familyname: Nguyen
  othernames: James Louis
  address: Monash University
  email: jlngu6@student.monash.edu
  correspondingauthor: true
  qualifications:  XXX
- familyname: Effendy
  othernames: Jimmy
  address: Monash University
  email: jeff0001@student.monash.edu
  correspondingauthor: true
  qualifications:  B.Com
- familyname: XXX
  othernames: XXX
  address: Monash University
  email: XXX
  correspondingauthor: true
  qualifications:  XXX
department: Econometrics and Business Statistics \newline add names &\newline add names
organization: Monash University
bibliography: references.bib
biblio-style: authoryear-comp
linestretch: 1.5
output:
  bookdown::pdf_document2:
    template: monashreport2.tex
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
    toc: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r libraries, include=FALSE}
library(tidyverse)
library(kableExtra)
library(bookdown)
library(lubridate)
```


```{r data-read, include=FALSE}
data = read.csv("data/wdi_data.csv")
data = data[1:35,]

names(data)[5:54] = c(1970:2019)
names(data)[3] = c("Series")

data_value = data %>%
  select('1970':'2019')

data_value = data.frame(lapply(data_value, as.character), stringsAsFactors =FALSE)
data_value = data.frame(lapply(data_value, as.numeric))

data = data.frame(Series = data$Series, data_value)
names(data)[2:51] = c(1970:2019)

data_long = data %>%
  pivot_longer(`1970`:`2019`, names_to = "Year") %>%
  mutate(Year = as.factor(Year))

#use data_long for analysis.
```


\section*{Introduction}

\section*{Motivation}

\section*{}

\section*{}

\section*{Education}

```{r education-expenditure}
data_education <- data_long %>% 
  filter(Series %in% (c("Ratio of female to male labor force participation rate (%) (national estimate)",
                        "Adjusted savings: education expenditure (current US$)",
                        "Government expenditure on education, total (% of GDP)",
                        "Adjusted savings: education expenditure (% of GNI)")))

#adding decades information to the dataframe
data_education$Year <- as.character(data_education$Year)
data_education$Year <- as.numeric(data_education$Year)

data_education <- data_education %>% mutate(decade = floor(Year/10)*10)

```

```{r ratio-plot-decade, fig.cap="Ratio of female to male labor force participation rate (%) (national estimate) by decades"}

data_education %>% filter(Series == "Ratio of female to male labor force participation rate (%) (national estimate)") %>%
  group_by(decade) %>% 
  summarise(average = mean(value, na.rm = TRUE)) %>% 
  ggplot(aes(x = decade,
             y = average)) +
  geom_col()
```

In the past 50 years, women's participation in the work force has increased steadily [@stier2000women].

The same trend can also be found in Israel - figure \@ref(fig:ratio-plot-decade) shows that the ratio of female to male labour force participate rate in Israel increases from approximately 50% in the 1970s to 85% in 2010s on average.

It is believed that gender inequality in education is a major contributing factor in gender inequalities in adulthood, especially in the labour market [@marks2008accounting].

In this section, the report is going to examine the association between education and the increase of women's employment in Israel.

```{r labor-edu-scatterplot, fig.cap = "Scatter Plot between Ratio of Female to Male Labour Force Participation Rate and Education Expenditure"}

data_education %>% filter(Series %in% (c("Adjusted savings: education expenditure (current US$)",
                                        "Ratio of female to male labor force participation rate (%) (national estimate)"))) %>%
  pivot_wider(names_from = Series,
              values_from = value) %>% 
  ggplot(aes(x = `Adjusted savings: education expenditure (current US$)`,
             y = `Ratio of female to male labor force participation rate (%) (national estimate)`,
             label = Year)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme(aspect.ratio = 1)

```

\subsection*{Education in Israel}

```{r total-edu-spending, fig.cap="Total Education Expenditure (US$) in Israel by decades"}
data_education %>% filter(Series == "Adjusted savings: education expenditure (current US$)") %>%
  #summarising education expenditure by decades
  group_by(decade) %>%
  summarise(avg_spending = mean(value, na.rm = TRUE)) %>% 
  ggplot(aes(x = decade,
             y = avg_spending)) +
  geom_col()
```

\section*{}

\section*{Model}

\section*{Conclusion}