---
title: "ETC5513 Project"
author:
- familyname: XXX
  othernames: XXX
  address: Monash University
  email: XXX
  correspondingauthor: true
  qualifications:  XXX
- familyname: Nguyen
  othernames: James Louis
  address: Monash University
  email: jlngu6@student.monash.edu
  correspondingauthor: true
  qualifications:  XXX
- familyname: XXX
  othernames: XXX
  address: Monash University
  email: XXX
  correspondingauthor: true
  qualifications:  XXX
- familyname: Tirodkar
  othernames: Siddhant
  address: Monash University
  email: stir0003@student.monash.edu
  correspondingauthor: true
  qualifications:  XXX
department: Econometrics and Business Statistics \newline add names &\newline add names
organization: Monash University
bibliography: references.bib
biblio-style: authoryear-comp
linestretch: 1.5
output:
  bookdown::pdf_document2:
    template: monashreport2.tex
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
    toc: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, include=FALSE}
library(tidyverse)
library(kableExtra)
library(bookdown)
library(lubridate)
```


```{r data-read, include=FALSE}
data = read.csv("data/wdi_data.csv")
data = data[1:35,]

names(data)[5:54] = c(1970:2019)
names(data)[3] = c("Series")

data_value = data %>%
  select('1970':'2019')

data_value = data.frame(lapply(data_value, as.character), stringsAsFactors =FALSE)
data_value = data.frame(lapply(data_value, as.numeric))

data = data.frame(Series = data$Series, data_value)
names(data)[2:51] = c(1970:2019)

data_long = data %>%
  pivot_longer(`1970`:`2019`, names_to = "Year") %>%
  mutate(Year = as.factor(Year))

#use data_long for analysis.
```


\section*{Introduction}

\section*{Motivation}

\section*{}

\section*{}

\section*{}

\section*{Labour Force Analysis}

This section of the report will focus on analysing various aspects of the Israelian labour force. The Israeli labour force has come a long way from being male dominated to being almost equally divided between men and women. The transition of the Israeli labour force shall be explored through the years 1991-2019 in this section of the report.
```{r labour-data}
israel_labour <- data_long %>% dplyr::filter(Series %in% c("Labor force, female (% of total labor force)",
                                                      "Ratio of female to male labor force participation rate (%) (national estimate)", "Self-employed, female (% of female employment) (modeled ILO estimate)" , "Wage and salaried workers, female (% of female employment) (modeled ILO estimate)",
                                                      "Labor force participation rate, female (% of female population ages 15-64) (modeled ILO estimate)","Labor force participation rate, male (% of male population ages 15-64) (modeled ILO estimate)"),
                                        Year %in% c(1991:2019))

```
\section*{Model}

\section*{Conclusion}