---
title: "ETC5513 Project"
author:
- familyname: XXX
  othernames: XXX
  address: Monash University
  email: XXX
  correspondingauthor: true
  qualifications:  XXX
- familyname: Nguyen
  othernames: James Louis
  address: Monash University
  email: jlngu6@student.monash.edu
  correspondingauthor: true
  qualifications:  XXX
- familyname: Lyubic
  othernames: Samuel
  address: Monash University
  email: slyu0007@student.monash.edu
  correspondingauthor: true
  qualifications:  Masters
- familyname: XXX
  othernames: XXX
  address: Monash University
  email: XXX
  correspondingauthor: true
  qualifications:  XXX
department: Econometrics and Business Statistics \newline add names &\newline add names
organization: Monash University
bibliography: references.bib
biblio-style: authoryear-comp
linestretch: 1.5
output:
  bookdown::pdf_document2:
    template: monashreport2.tex
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
    toc: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, include=FALSE}
library(tidyverse)
library(kableExtra)
library(bookdown)
library(lubridate)
library(gridExtra)
```


```{r data-read, include=FALSE}
data = read.csv("data/wdi_data.csv")
data = data[1:36,]

names(data)[5:54] = c(1970:2019)
names(data)[3] = c("Series")

data_value = data %>%
  select('1970':'2019')

data_value = data.frame(lapply(data_value, as.character), stringsAsFactors =FALSE)
data_value = data.frame(lapply(data_value, as.numeric))

data = data.frame(Series = data$Series, data_value)
names(data)[2:51] = c(1970:2019)

data_long = data %>%
  pivot_longer(`1970`:`2019`, names_to = "Year") %>%
  mutate(Year = as.factor(Year))

#use data_long for analysis.
```


\section*{Introduction}

\section*{Motivation}

\section*{}

\section*{}

\section*{Male and Female Mortality Rates and Birth Rate}

With Israel as our chosen country, this section will be analyzing the relationship between our repose variable, _Ratio of female to male labor force participation rate (%) (national estimate)_, and the following economic indicators: _Birth rate, crude (per 1,000 people)_, _Mortality rate, adult, female (per 1,000 female adults)_ and _Mortality rate, adult, male (per 1,000 male adults)_ to assess their impact on the change of the female to male labor force ratio to ultimately address the change in how the gender stereotype is viewing in Israel. 
Across history, the traditional stereotype on gender roles has been one that depicts males roles as the provider and women as the caretaker and child bearer resulting in the notion that a woman's role is at home which has led to less occupational opportunities and aspirations \cite{GenderRoles}. These stereotypes are further perpetuated through religion, as seen by the two most present religion in Israel - Judaism and Islam where the traditional identity for involves leadership for men and domesticity for females \cite{Woodhead} in all facets of life from work, private, public and in the army \cite{ServiceIDF}. However, it does seem as though this traditional view is being confronted and dismantled over time with more religious and non religious Israeli woman taking up education and being accepted in a range of occupational opportunities and army duties \cite{Levy2006}, as such birth rate and male and female mortality rates are prime variables to potentially explain the change in the labor force ratio thus indicating the potential change in how gender stereotypes are viewed. 

The variables for this section are isolated into their own data set in both long and wide form.
```{r create-long-wide-objects-with-required-variables}
mortality_birth_rate_long <- data_long %>%
  filter(Series %in% c("Ratio of female to male labor force participation rate (%) (national estimate)", "Mortality rate, adult, female (per 1,000 female adults)", "Mortality rate, adult, male (per 1,000 male adults)", "Birth rate, crude (per 1,000 people)")) %>%
  arrange(Year) %>%
  slice(65:188)
```

### The variable change overtime

```{r figure-female-and-male-mortality-and-birth-rate-and-labor-force-ratio, include = FALSE}
figure_female_male_mortality <- mortality_birth_rate_long %>%
  filter(Series %in% c("Mortality rate, adult, female (per 1,000 female adults)", 
                       "Mortality rate, adult, male (per 1,000 male adults)")) %>%
  ggplot(aes(x = Year,
             y = value,
             group = Series,
             color = Series)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "right") +
  theme_bw() +
  labs(title = "Male and female mortality rate, per 1000 adults",
       y = "Mortality rate per 1000 people",
       x = "Year") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) 

figure_birth_rate <- mortality_birth_rate_long %>%
  filter(Series == "Birth rate, crude (per 1,000 people)") %>%
  ggplot(aes(x = Year,
             y = value,
             group = Series,
             color = Series)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "right") +
  theme_bw() +
  labs(title = "Birth rate per 1000 people",
       y = "Birth rate, crude (per 1,000 people)",
       x = "Year") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

figure_labor_force_ratio <- mortality_birth_rate_long %>%
  filter(Series == "Ratio of female to male labor force participation rate (%) (national estimate)") %>%
  ggplot(aes(x = Year,
             y = value,
             group = Series,
             color = Series)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "right") +
  theme_bw() +
  labs(title = "Ratio female to male labor force participation",
       y = "Female to male ratio %",
       x = "Year") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) 
  
```

```{r combining-the-variable-figures, fig.cap = "Each figure visualises the chosen variables", fig.width = 10, fig.height = 8}
figure_combined_birth_ratio_mortality <- grid.arrange(figure_labor_force_ratio, figure_birth_rate, figure_female_male_mortality)
```

```{r percent-change-over-the-observed-time-period, fig.cap = "Figure showing the percentage chance in each of the chosen variables"}
pct_change_mortality_variables <- mortality_birth_rate_long %>%
  filter(Year %in% c("1986", "2016")) %>% 
  pivot_wider(names_from = Year,
              values_from = value) %>%
  mutate(percent_change_1986_to_2016 = ((`2016`-`1986`)/`1986`)*100) %>%
  kable(caption = "The total percentage change from 1986 to 2016, per each variable") %>%
  kable_styling(bootstrap_options = c("hover", "striped"),
                latex_options = "hold_position") 
pct_change_mortality_variables    
```

Figure \@ref(fig:combining-the-variable-figures) visualizes the change over time for each of the chosen variables in this section, with the selected variable on the y axis and year along the x axis. 

Key takeaways:
  - The labor force participation figure shows that the ratio of female to male labor force participation rate has increased from 1986 to 2016, by 41.01% as shown by Table \@ref(tab:pct_change_mortality_variables).
  - The birth rate figure shows the birth rate per 1000 people overall shows a reduction in the birth rate has increased from 1986 to 2016, by -8.23% as shown by Table \@ref(tab:pct_change_mortality_variables).
  - The mortality rate table shows the mortality rate for females per 1000 female adults in orange and the mortality rate for males per 1000 males in blue. Both indicators have reduced from 1986 to 2016 at a relatively even rate, with female mortality reducing by -48.94% and male mortality reducing by -45.80%, as shown by Table \@ref(tab:pct_change_mortality_variables).

### Correlation comparison

The explanatory variables are presented as ratios and in order for their results to be relative to the explanatory variable and to allow for analysis to take place, they are turned into percentages. The new column percent_value shows the percentage of birth rate and female and male mortality rates for every 1000 people, from which a new data set will be created showing the percentage results for all the variables in this section.
```{r ratios-turned-into-percentages, include = FALSE}
variables_turned_into_percentage <- mortality_birth_rate_long %>%
  filter(!Series == "Ratio of female to male labor force participation rate (%) (national estimate)") %>%
  mutate(percent_value = ((value/1000)*100))
```

```{r combine-percent-results-response-variable-and-explanatory, include = FALSE}
labor_force_ratio <- mortality_birth_rate_long %>%
  filter(Series == "Ratio of female to male labor force participation rate (%) (national estimate)") %>%
  mutate(percent_value = value)
percentage_results_all_variables <- rbind(variables_turned_into_percentage, labor_force_ratio)
```

```{r pivot-wider-percentage-results, include = FALSE}
percentage_results_wider <- percentage_results_all_variables %>% 
  select(Series, Year, percent_value) %>%
  pivot_wider(names_from = Series,
              values_from = percent_value) %>%  
  rename(Pct_male_mortality_rate_per_1000_male_adults = "Mortality rate, adult, male (per 1,000 male adults)",
         Pct_Female_mortality_rate_1000_female_adults = "Mortality rate, adult, female (per 1,000 female adults)",
         Pct_Birth_rate_per_1000_people = "Birth rate, crude (per 1,000 people)",
         Labor_force_rate = "Ratio of female to male labor force participation rate (%) (national estimate)")
```

```{r correlation-between-response-explanatory, fig.cap = "The table shows the correlation coefficient between each explanatory variable and the response variable"}
correlation_results_male_mort <- cor(percentage_results_wider$Pct_male_mortality_rate_per_1000_male_adults, percentage_results_wider$Labor_force_rate)
correlation_results_female_mort <- cor(percentage_results_wider$Pct_Female_mortality_rate_1000_female_adults, percentage_results_wider$Labor_force_rate) 
correlation_results_birth_rate <- cor(percentage_results_wider$Pct_Birth_rate_per_1000_people, percentage_results_wider$Labor_force_rate)

table_correlation <- tibble(correlation_results_male_mort, correlation_results_female_mort, correlation_results_birth_rate) %>%
  kable(caption = "The correlation between the percentage results, per 1000 people, of the response variables and the ratio of female to male labor force participation rate (%)") %>%
  kable_styling(bootstrap_options = c("hover", "striped"),
                latex_options = "hold_position") 
table_correlation
```

```{r figure-correlation-female-mort-labor-force, fig.cap = "The figure displays the relationship between the percentage results of female mortality rates per 1000 adult females and the Ratio of female to male labor force participation rate (%) (national estimate)"}
correlation_pct_results_female_mort <- percentage_results_wider %>%
  ggplot(aes(x = Labor_force_rate, 
             y = Pct_Female_mortality_rate_1000_female_adults)) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE) +
  labs(title = "Correlation between the female mortality 
                rate and labor force ratio",
       y = "Female mortality per 1000 adult females, as a percentage",
       x = "Ratio female to male labor force participation")
correlation_pct_results_female_mort
```

```{r figure-correlation-male-mort-labor-force, fig.cap = "The figure displays the relationship between the percentage results of male mortality rates per 1000 adult females and the Ratio of female to male labor force participation rate (%) (national estimate)"}
correlation_pct_results_male_mort <- percentage_results_wider %>%
  ggplot(aes(x = Labor_force_rate, 
             y = Pct_male_mortality_rate_per_1000_male_adults)) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE) +
  labs(title = "Correlation between the male mortality 
                rate and labor force ratio",
       y = "Male mortality per 1000 adult males, as a percentage",
       x = "Ratio female to male labor force participation")
correlation_pct_results_male_mort
```

```{r figure-correlation-birth-rate-labor-force, fig.cap = "The figure displays the relationship between the percentage results of birth rates per 1000 adult females and the Ratio of female to male labor force participation rate (%) (national estimate)", echo = FALSE}
correlation_pct_results_birth_rate <- percentage_results_wider %>%
  ggplot(aes(x = Labor_force_rate, 
             y = Pct_Birth_rate_per_1000_people)) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE) +
  labs(title = "Correlation between birth rate and labor force ratio",
       y = "Birth rate per 1000 people, as a percentage",
       x = "Ratio female to male labor force participation")
correlation_pct_results_birth_rate
```

Figure \@ref(fig:figure-correlation-birth-rate-labor-force) visualizes the correlation between birth rate and the female to male labor force ratio, with birth rate, crude (per 1,000 people) on the y axis and the ratio of female to male labor force participation rate on the x axis. The figure shows a negative correlation between the two variables however, there are some scattered results where both birth rate and female to male labor force ratio are both relatively low which is depicted by the correlation coefficient of -0.60 as shown by Table \@ref(tab:correlation-between-response-explanatory).


Figure \@ref(fig:figure-correlation-male-mort-labor-force) visualizes the correlation between male mortality rate and the female to male labor force ratio, with male mortality rate per 1,000 male adults on the y axis and the ratio of female to male labor force participation rate on the x axis. The figure displays a strong negative correlation between the two variables which is backed up by the correlation coefficient of -.96 in Table \@ref(tab:correlation-between-response-explanatory). 

Figure \@ref(fig:figure-correlation-female-mort-labor-force) visualizes the correlation between female mortality rate and the female to male labor force ratio, with female mortality rate per 1,000 female adults on the y axis and the ratio of female to male labor force participation rate on the x axis. The figure displays a strong negative correlation between the two variables which is backed up by the correlation coefficient of -.99 in Table \@ref(tab:correlation-between-response-explanatory).

As evident by the analysis conducted in this section, there is a clear positive correlation between the three explanatory variables and the the ratio of female to male labor force participation rate.  



\section*{}

\section*{Model}

\section*{Conclusion}

references