---
title: "ETC5513 Project"
author:
- familyname: XXX
  othernames: XXX
  address: Monash University
  email: XXX
  correspondingauthor: true
  qualifications:  XXX
- familyname: Nguyen
  othernames: James Louis
  address: Monash University
  email: jlngu6@student.monash.edu
  correspondingauthor: true
  qualifications:  XXX
- familyname: Lyubic
  othernames: Samuel
  address: Monash University
  email: slyu0007@student.monash.edu
  correspondingauthor: true
  qualifications:  Masters
- familyname: XXX
  othernames: XXX
  address: Monash University
  email: XXX
  correspondingauthor: true
  qualifications:  XXX
department: Econometrics and Business Statistics \newline add names &\newline add names
organization: Monash University
bibliography: references.bib
biblio-style: authoryear-comp
linestretch: 1.5
output:
  bookdown::pdf_document2:
    template: monashreport2.tex
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
    toc: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, include=FALSE}
library(tidyverse)
library(kableExtra)
library(bookdown)
library(lubridate)
```


```{r data-read, include=FALSE}
data = read.csv("data/wdi_data.csv")
data = data[1:36,]

names(data)[5:54] = c(1970:2019)
names(data)[3] = c("Series")

data_value = data %>%
  select('1970':'2019')

data_value = data.frame(lapply(data_value, as.character), stringsAsFactors =FALSE)
data_value = data.frame(lapply(data_value, as.numeric))

data = data.frame(Series = data$Series, data_value)
names(data)[2:51] = c(1970:2019)

data_long = data %>%
  pivot_longer(`1970`:`2019`, names_to = "Year") %>%
  mutate(Year = as.factor(Year))

#use data_long for analysis.
```


\section*{Introduction}

\section*{Motivation}

\section*{}

\section*{Electricity generation distribution}

```{r}
data_electric = data_long %>%
  filter(Series %in% c("Renewable electricity output (% of total electricity output)",
                       "Fossil fuel energy consumption (% of total)",
                       "Energy imports, net (% of energy use)",
                       "Fossil fuel energy consumption (% of total)",
                       "Renewable electricity output (% of total electricity output)",
                       "Electricity production from oil, gas and coal sources (% of total)"))
```

```{r}
data_electric %>% 
  filter(Series %in% c("Fossil fuel energy consumption (% of total)")) %>%
  drop_na() %>%
  ggplot(aes(x=Year, y=value, colour=Series)) + 
  geom_line(aes(group=1)) + 
  theme(axis.text.x = element_text(angle = 90))

data_electric %>% 
  filter(Series %in% c("Energy imports, net (% of energy use)")) %>%
  drop_na() %>%
  ggplot(aes(x=Year, y=value, colour=Series)) + 
  geom_line(aes(group=1)) + 
  theme(axis.text.x = element_text(angle = 90))

data_electric %>% 
  filter(Series %in% c("Renewable electricity output (% of total electricity output)")) %>%
  drop_na() %>%
  ggplot(aes(x=Year, y=value, colour=Series)) + 
  geom_line(aes(group=1)) + 
  theme(axis.text.x = element_text(angle = 90))
```

** Graph of the energy production compare.
** Primary focus on the renewables.

** Table of rate of change, plot this rate of change
** Use average of rate of change to see if they are able to get to that level predicted.

https://iea-pvps.org/wp-content/uploads/2020/05/IEA-PVPS-AR-2019-1.pdf

In 2019, Israel reached a level ~ 7% of total energy production from renewables.
70% of total was natural gas, I am guessing that the rest was from coal.

Changing attitudes - target 17% by 2030, interim 10% by 2020.

Renewable energy is produced by solar fields. 95% of the above 7% was solar energy (renewable sources).

Another point about changing attitudes, plan was initated in 2016.

Israeli network is island * no electrical grid power.

Abundant natural gas in israel, after 2000 (first commercial discovery of gas in israel), became self-sufficient/net exporter of gas.

very focussed to change. https://www.export.gov/apex/article2?id=Israel-En


\section*{Male and Female Mortality Rates}

\section*{}

\section*{Model}

\section*{Conclusion}