---
title: "ETC5513 Project"
author:
- familyname: XXX
  othernames: XXX
  address: Monash University
  email: XXX
  correspondingauthor: true
  qualifications:  XXX
- familyname: Nguyen
  othernames: James Louis
  address: Monash University
  email: jlngu6@student.monash.edu
  correspondingauthor: true
  qualifications:  XXX
- familyname: Lyubic
  othernames: Samuel
  address: Monash University
  email: slyu0007@student.monash.edu
  correspondingauthor: true
  qualifications:  Masters
- familyname: XXX
  othernames: XXX
  address: Monash University
  email: XXX
  correspondingauthor: true
  qualifications:  XXX
department: Econometrics and Business Statistics \newline add names &\newline add names
organization: Monash University
bibliography: references.bib
biblio-style: authoryear-comp
linestretch: 1.5
output:
  bookdown::pdf_document2:
    template: monashreport2.tex
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    includes:
      in_header: preamble.tex
    keep_tex: yes
    number_sections: yes
    citation_package: biblatex
    toc: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, include=FALSE}
library(tidyverse)
library(kableExtra)
library(bookdown)
library(lubridate)
```


```{r data-read, include=FALSE}
data = read.csv("data/wdi_data.csv")
data = data[1:36,]

names(data)[5:54] = c(1970:2019)
names(data)[3] = c("Series")

data_value = data %>%
  select('1970':'2019')

data_value = data.frame(lapply(data_value, as.character), stringsAsFactors =FALSE)
data_value = data.frame(lapply(data_value, as.numeric))

data = data.frame(Series = data$Series, data_value)
names(data)[2:51] = c(1970:2019)

data_long = data %>%
  pivot_longer(`1970`:`2019`, names_to = "Year") %>%
  mutate(Year = as.factor(Year))

#use data_long for analysis.
```


\section*{Introduction}

\section*{Motivation}

\section*{}

\section*{Renewable energy}

Israel has seen significant changes in the distribution of genders in the labour force. Though many direct effects can be explored, this section looks at changes in attitude by analysing Israel's move to more renewable sources of energy. There may be a link between the a transition to change in political thinking in priority. This may also be attributed to a general change in attitude, and could provide a reasonable proxy for our model.

For this section of the report, we will be looking at World Bank Sustainable for All data. This data set comprises of electricity data between 1990 and 2015. 
```{r}
data_sea = read.csv("data/SEA_data.csv")
names(data_sea)=c("Country Name", "Country Code", "Year", "Year Code",
                  "Renewable electricity share of total electricity output (%)",
                  "Renewable energy consumption (TJ)",
                  "Total final energy consumption (TFEC) (TJ)",
                  "Renewable electricity output (GWh)",
                  "Renewable energy share of TFEC (%)")
data_sea = data_sea[1:26,]

data_sea_long = data_sea %>%
  pivot_longer(`Renewable electricity share of total electricity output (%)`:`Renewable energy share of TFEC (%)`,names_to = "Series")
```

```{r}
data_sea_long %>% 
  filter(Series == "Renewable electricity share of total electricity output (%)") %>%
  ggplot(aes(x = Year, y = value, colour = Series)) + 
  geom_line(aes(group = 1))
```



```{r}
data_electric = data_long %>%
  filter(Series %in% c("Renewable electricity output (% of total electricity output)",
                       "Fossil fuel energy consumption (% of total)",
                       "Energy imports, net (% of energy use)",
                       "Fossil fuel energy consumption (% of total)",
                       "Renewable electricity output (% of total electricity output)",
                       "Electricity production from oil, gas and coal sources (% of total)",
                       "Electricity production from renewable sources, excluding hydroelectric (% of total)"))
```

```{r}
data_electric %>% 
  filter(Series %in% c("Fossil fuel energy consumption (% of total)")) %>%
  drop_na() %>%
  ggplot(aes(x=Year, y=value, colour=Series)) + 
  geom_line(aes(group=1)) + 
  theme(axis.text.x = element_text(angle = 90))

data_electric %>% 
  filter(Series %in% c("Energy imports, net (% of energy use)")) %>%
  drop_na() %>%
  ggplot(aes(x=Year, y=value, colour=Series)) + 
  geom_line(aes(group=1)) + 
  theme(axis.text.x = element_text(angle = 90))

data_electric %>% 
  filter(Series %in% c("Electricity production from renewable sources, excluding hydroelectric (% of total)")) %>%
  drop_na() %>%
  ggplot(aes(x=Year, y=value, colour=Series)) + 
  geom_line(aes(group=1)) + 
  theme(axis.text.x = element_text(angle = 90)) + 
  ylab("Percentage of total electricity output (%)")
```

** Graph of the energy production compare.
** Primary focus on the renewables.

** Table of rate of change, plot this rate of change
** Use average of rate of change to see if they are able to get to that level predicted.

https://iea-pvps.org/wp-content/uploads/2020/05/IEA-PVPS-AR-2019-1.pdf

In 2019, Israel reached a level ~ 7% of total energy production from renewables.
70% of total was natural gas, I am guessing that the rest was from coal.

Changing attitudes - target 17% by 2030, interim 10% by 2020.

Renewable energy is produced by solar fields. 95% of the above 7% was solar energy (renewable sources).

Another point about changing attitudes, plan was initated in 2016.

Israeli network is island * no electrical grid power.

Abundant natural gas in israel, after 2000 (first commercial discovery of gas in israel), became self-sufficient/net exporter of gas.

very focussed to change. https://www.export.gov/apex/article2?id=Israel-En


\section*{Male and Female Mortality Rates}

\section*{}

\section*{Model}

\section*{Conclusion}